---
title: "Exercise 5"
author: "Carilli"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
library(broom)
library(ggthemes)
library(magrittr)
library(kableExtra)
```

```{r 5-data}
height_wage <- read_csv("heightwage_british_males.csv")
```

### (a)

```{r 5a}
height_wage %>%
  filter(gwage33 < 400 & height33 > 40) %>% 
  lm(gwage33 ~ height33, .) %>% 
  jtools::summ() 
```  

```{r 5a-df, echo=FALSE}
df <- 
height_wage %>%
  filter(gwage33 < 400 & height33 > 40) %$% 
  lm(gwage33 ~ height33) %>% 
  glance() %>% 
  pull(df.residual)

tidy_stats <- 
height_wage %>%
  filter(gwage33 < 400 & height33 > 40) %$% 
  lm(gwage33 ~ height33) %>% 
  tidy()
```

Given, the critical *t*-statistic at the $\alpha=5\%$ level of significance with `r df` degrees of freedom is `r qt(.975, df)`, we fail to reject $H_0:\beta_0=0$ and reject $H_1:\beta_1=0$.

### (b)

The *p* values indicate the probability that we would see the value of each coefficient if $H_0$ were true.  For $\hat{\beta_0}$, there is a `r tidy_stats$p.value[1]` probability that we would observe a value of `r tidy_stats$estimate[1]` if $\beta_0$ were actually zero. This is not a small enough probability to give us evidence to suggest that $\beta_0\ne0$. For $\hat{\beta_1}$, there is a `r tidy_stats$p.value[2]` probability that we would observe a value of `r tidy_stats$estimate[2]` if $\beta_1$ were actually zero. This is a large enough probability to give us evidence to suggest that $\beta_1\ne0$.

### (c)

```{r 5c, echo=FALSE}
conf_int <- 
height_wage %>%
  filter(gwage33 < 400 & height33 > 40) %$% 
  lm(gwage33 ~ height33) %>% 
  tidy() %>% 
  select(beta = estimate, se = std.error, t = statistic)
```

The confidence interval is calculate as $\hat{\beta_i} \pm t_{n-2}se_{\hat{\beta_i}}$.  The confidence interval for $\beta_0$ is
$$\begin{align}
`r conf_int$beta[1]` &\pm `r qt(.975, df)` \times `r conf_int$se[1]`\\
&\pm `r qt(.975, df)*conf_int$se[1]`\\
`r conf_int$beta[1] - qt(.975, df) * conf_int$se[1]` \le &\beta_0 \le `r conf_int$beta[1] + qt(.975, df) * conf_int$se[1]` 
\end{align}$$
The confidence interval for $\beta_1$ is
$$\begin{align}
`r conf_int$beta[2]` &\pm `r qt(.975, df)` \times `r conf_int$se[2]`\\
&\pm `r qt(.975, df)*conf_int$se[2]`\\
`r conf_int$beta[2] - qt(.975, df) * conf_int$se[2]` \le &\beta_1 \le `r conf_int$beta[2] + qt(.975, df) * conf_int$se[2]` 
\end{align}$$

Alternatively we could add the argument `conf.int = TRUE` to `broom::tidy()`

```{r 5c-alt}
height_wage %>%
  filter(gwage33 < 400 & height33 > 40) %$% 
  lm(gwage33 ~ height33) %>% 
  tidy(conf.int  = TRUE)
```

### (d)

Given that the *p* value reported in part (a) above is `r tidy_stats$p.value[2]`. We would reject $H_0:\beta_1=0$.

### (e)

Given that we failed to reject $H_0:\beta=0$ at the the $\alpha=5\%$ level of significance, we would also fail to reject $H_0$ at the $\alpha=1\%$ level of significance.

### (f)

```{r 5f}
height_wage %>% 
  slice(1:800) %>% 
  lm(gwage33 ~ height33, .) %>% 
  tidy() 
```

We fail to reject $H_0:\beta_i=0 \hspace{.5cm} \forall i$.
