---
title: "Exercise 3"
author: "Carilli"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
require(tidyverse)
require(broom)
require(magrittr)
require(kableExtra)
require(forcats)
require(moderndive)
```

```{r data}
pres_vote <- read_csv("PresPartyEconGrowth1.csv")
```

### (a)

```{r part_a}
pres_vote %$% 
  lm(unemployment ~ lagdempresident) %>%
  tidy() %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

$\hat{\beta_1}$ tells us that if last year's president was a democrat we'd expect the unemployment rate to be 0.955 percentage points lower.  $\hat\beta_0$ tells us that if the president last year was a democrat that the unemployment rate is 6.24%. 

### (b)

```{r part_b}
pres_vote %$% 
  lm(changegdppc ~ lagdempresident) %>%
  tidy() %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

$\hat{\beta_1}$ tells us that if last year's president was a democrat we'd expect real GDP per capita to be \$220 higher.  $\hat\beta_0$ tells us that if the president last year was a democrat that real GDP per capita would be $481.

### (c)

Let $\alpha = 5\%$, we fail to reject the $H_0: \beta_1=0$ since the *p* value is greater than $\alpha$ in each question.

### (d)

The *p* value indicates the probability of creating a type I error, that is of wrongly rejecting $H_0$.

### (e)

```{r part_e}
std_error_beta_1 <- 
pres_vote %$% 
  lm(changegdppc ~ lagdempresident) %>% 
  tidy() %>% 
  filter(term == "lagdempresident") %>% 
  pull(std.error)
beta_range <- seq(0, 1000, 25)
power_curve <- pnorm(beta_range/std_error_beta_1 - qnorm(.99))
tibble(beta_range, power_curve) %>% 
  ggplot(aes(x = beta_range, power_curve)) + 
  geom_line() + 
  xlab("Beta") +
  ylab("Probability reject the null")


```

