---
title: "Exercise 3"
author: "Carilli"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: cerulean
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, warning = FALSE, message = FALSE,
                      options(digits = 3, scipen = 999))
library(tidyverse)
```

```{r 3-data}
pres_vote <- 
  here::here("data", "PresPartyEconGrowth1.csv") %>% 
  read_csv() %>% 
  janitor::clean_names()
```

### (a)

```{r 3a}
pres_vote %>% 
  lm(unemployment ~ lagdempresident, .) %>%
  jtools::summ()
```

```{r 3a2, echo=FALSE}
pres_vote %>% 
  lm(unemployment ~ lagdempresident, .) %>% 
  tidy() %>% 
  pull(estimate) -> beta_hats
```

$\hat{\beta_1}$ tells us that if last year's president was a democrat we'd expect the unemployment rate to be `r beta_hats[2]` percentage points lower.  $\hat\beta_0$ tells us that if the president last year was a democrat that the unemployment rate was `r beta_hats[1]`%. 

### (b)

```{r 3b}
pres_vote %>% 
  lm(changegdppc ~ lagdempresident, .) %>%
  jtools::summ()
```

```{r 3b2, echo=FALSE}
pres_vote %>% 
  lm(changegdppc ~ lagdempresident, .) %>%
  tidy() %>% 
  pull(estimate) -> beta_hats
```


$\hat{\beta_1}$ tells us that if last year's president was a democrat we'd expect real GDP per capita to be \$`r beta_hats[2]` higher.  $\hat\beta_0$ tells us that if the president last year was a democrat that real GDP per capita would be \$`r beta_hats[1]`.

### (c)

Let $\alpha = 5\%$, we fail to reject the $H_0: \beta_1=0$ since the *p* value is greater than $\alpha$ in each question.

### (d)

The *p* value indicates the probability of creating a type I error, that is of wrongly rejecting $H_0$.

### (e)

```{r 3e}
std_error_beta_1 <-
  pres_vote %>%
  lm(changegdppc ~ lagdempresident, .) %>%
  broom::tidy() %>%
  filter(term == "lagdempresident") %>%
  pull(std.error)
beta_range <- seq(0, 1000, 25)
power_curve <- pnorm(beta_range / std_error_beta_1 - qnorm(.99))
tibble(beta_range, power_curve) %>%
  ggplot(aes(x = beta_range, power_curve)) +
  geom_line() +
  xlab(expression(beta[1])) +
  ylab(expression(paste("Probability of rejecting ", H[0])))


```





```{r}
jtools::export_summs(
  pres_vote %>% 
  lm(unemployment ~ lagdempresident, .),
  
  pres_vote %>% 
  lm(changegdppc ~ lagdempresident, .)
)


```

